### 2018-03-14 paper/3-abstractions.tex
Finish S3.3 and S3.4

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store  = initialise () (\key -> if key == "A1" then 10 else 20)
@\ghci@ result = busy sprsh1 "B2" store
@\ghci@ getValue result "B1"
30
@\ghci@ getValue result "B2"
60
\end{minted}
\vspace{1mm}

\noindent
As one can see, \hs{busy} correctly built both \cmd{B1} and \cmd{B2}, and in the
right order (if it had built \cmd{B2} before building \cmd{B1}, the result would
have been $20 * 2 = 40$ instead of $(10 + 20) * 2 = 60$). As an example showing
that \hs{busy} is not minimal, imagine that the formula in cell \cmd{B2} was
\cmd{B1~+~B1} instead of \cmd{B1~*~2}. This would lead to calling
\hs{compute}~\hs{"B1"} twice -- once per each occurrence of \cmd{B1} in the
formula.

An alert reader might notice that \hs{busy} accepts \emph{monadic tasks} of
type \hs{Task}~\hs{Monad}, whereas \hs{sprsh1} is a \hs{Task}~\hs{Applicative}.
This indicates that \hs{busy} can cope with tasks containing dynamic
dependencies, which we address in the next subsection~\S\ref{sec-task-monad}.

@@

As discussed in~\S\ref{sec-background-excel}, some build tasks have dynamic
dependencies, which are determined by values of intermediate computations. In
our framework, such build tasks correspond to the type
\hs{Task}~\hs{Monad}~\hs{k}~\hs{v}. Recall the following spreadsheet
example:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{text}
A1: 10      B1: IF(C1=1,A1,A2)      C1: 1
A2: 20
\end{minted}
\vspace{1mm}

@@

\noindent
Since the \hs{busy} build system introduced in~\S\ref{sec-general-build} always
rebuilds every dependency it encounters, it is easy for it to handle dynamic
dependencies.

\todo{AM}{Describe another simple build system that does topological sort and
builds everything in order. It will not be able to handle dynamic dependencies.}

>>>>

\begin{minted}[xleftmargin=10pt]{haskell}
@\ghci@ store  = initialise () (\key -> if key == "A1" then 10 else 20)
@\ghci@ result = busy sprsh1 "B2" store
@\ghci@ getValue result "B1"
30
@\ghci@ getValue result "B2"
60
\end{minted}

\noindent
As one can see, \hs{busy} built both \cmd{B2} and its dependency \cmd{B1} in the
right order (if it had built \cmd{B2} before building \cmd{B1}, the result would
have been $20 * 2 = 40$ instead of $(10 + 20) * 2 = 60$). As an example showing
that \hs{busy} is not minimal, imagine that the formula in cell \cmd{B2} was
\cmd{B1~+~B1} instead of \cmd{B1~*~2}. This would lead to calling
\hs{compute}~\hs{"B1"} twice -- once per each occurrence of \cmd{B1} in the
formula.

An alert reader might notice that \hs{busy} accepts \emph{monadic tasks} of
type \hs{Task}~\hs{Monad}, whereas \hs{sprsh1} is a \hs{Task}~\hs{Applicative}.
This indicates that \hs{busy} can cope with tasks containing dynamic
dependencies, which we discuss in the next subsection~\S\ref{sec-task-monad}.

It is interesting to note that if \hs{busy} is given a \emph{functorial task}
of type \hs{Task}~\hs{Functor}, which cannot use the application operator
\hs{<*>}, then \hs{busy} can rejoice in knowing that it will never have to
rebuild a key twice. Indeed, functorial tasks are very much like tail recursive
functions -- they cannot branch into multiple subproblems, always leading to a
straight-line build.

@@

As explained in~\S\ref{sec-background-excel}, some build tasks have dynamic
dependencies, which are determined by values of intermediate computations. In
our framework, such build tasks correspond to the type
\hs{Task}~\hs{Monad}~\hs{k}~\hs{v}. Recall the following spreadsheet
example:

\vspace{1mm}
\begin{minted}[xleftmargin=10pt]{text}
A1: 10      B1: IF(C1=1,A1,A2)      C1: 1
A2: 20
\end{minted}

@@

\noindent
The main difference compared to \hs{sprsh1} is that the computation now takes
place in a \hs{Monad}, which allows us to extract the value of \hs{c1} and
\hs{fetch} different keys depending on whether \hs{c1}~\hs{==}~\hs{1} or not.

Since the \hs{busy} build system introduced in~\S\ref{sec-general-build} always
rebuilds every dependency it encounters, it is easy for it to handle dynamic
dependencies. For minimal build systems, however, dynamic dependencies and hence
monadic tasks are much more challenging, as we will see
in~\S\ref{sec-implementations}.
